\sysfull

### 2017-11-15 paper/compiler.tex
review of compiler section

We studied \sys's effectiveness for systems with automatically generated task
boundaries by developing a simple compiler prototype that transforms an
arbitrary C into a program with tasks. The compiler is representative of,
although not a direct implementation of, a system that assists in writing
task-based programs by automatically inserting tasks (e.g., inspired by
Ratchet~\cite{ratchet}). 
%
Translating C code to use tasks requires identifying program points to use as
task boundaries, defining a set of task functions that begin and end at those
boundaries, ensuring that all existing control-flow paths are either contained
within a task or are made explicit with inter-task control-flow, and identifying
and instrumenting accesses to task-shared variables that need to be protected
by \sys at runtime. Our simple prototype, which is implemented using LLVM, analyzes the
control-flow graph (CFG) and uses a rule-based heuristic to determine at which
points to place a task boundary. After identifying where to put task
boundaries, the \sys compiler transforms the imperative code into tasks,
instrumenting protected memory accesses.

@@

\sys analyzes a program to determine where to put task boundaries. 
%
A task is single-entry, multiple-exit part of the CFG that our compiler excises
into a new function.  
%
There are several factors that determine whether tasks will execute
efficiently. A task decomposition should minimize the number of variables that
are accessed by multiple tasks and, consequently, must be protected by \sys's
paging mechanism. 
%
Different tasks should be approximately the same size to make reasoning about
coalescing simple by eliminating the need to consider variable run time costs. 
%
To minimize task boundary overheads (e.g., commit), a task decomposition should
maximize task size, which minimizes the frequency of task transitions.

@@

\begin{figure}
	\centering
	\includegraphics[width=.85\columnwidth]{figures/compiler.pdf}
	\caption{\sys compiles code to use tasks.}
	\label{fig:compiler_overview}
\end{figure}

The task boundary identification analysis processes a standard C program (as
shown in Figure~\ref{fig:compiler_overview}) and produces a CFG with some
vertices marked as {\em task boundaries}. The \sys compiler builds the
program's CFG as shown in Figure~\ref{fig:compiler_overview}. In our simple
reference prototype, we implemented inter-procedural analysis by inlining all
functions called.  We opted not to support recursion, which is uncommon in
embedded code. We leave to future work better approaches to interprocedural
analysis and task transformation.

@@

{\noindent \bf Estimated run time.} \sys assigns each basic block in the CFG a
{\em cost}, which is an estimate of the run time of the block.  Precisely
modeling the run time and energy cost of arbitrary C code is difficult and
beyond the scope of our simple compiler prototype.  Instead, our prototype counts
LLVM intermediate representation (IR) instructions and assigns a higher cost to
a block with more operations, assuming that the block runs for a longer period
of time and consumes more energy.  
%
To make the cost estimate useful, our compiler prototype estimates the maximum
cost that the device's energy buffering capacitor supports.  \sys estimates
this quantity by measuring the number of arithmetic operations that can execute
without a power interruption, starting from a full capacitor. This maximum number
of operations is a limit on the cost of a task -- any longer task risks not
executing to completion.

{\noindent \bf Live variable cost.} Each CFG edge is associated with a set of
variables that {\em may} be {\em alive} across the edge (i.e., variables
assigned before the edge and may be read after the edge). Variables that live
across an edge must be protected by \sys's memory virtualization mechanism when
\sys places a task boundary on the edge.  \sys heuristically minimizes variable
virtualization overhead by placing boundaries to minimize the number of
live variables that span task boundaries.

{\noindent \bf Loop cost.} Loops are a challenge for \sys because the analysis
must account for the cost of multiple iterations of the loop body. \sys assumes
that most loops are annotated with a static loop bound. For loops that do not
have a static bound, \sys conservatively assumes the loop will be too large to
fit within a single task. Assuming loop bounds is a reasonable simplification:
while not always the case, long loops in embedded code are often statically
bounded, especially when code aims to meet strict resource requirements.

@@

After placing task boundaries, the compiler must ensure that all tasks are
single-entry. To do so, \sys computes the dominator tree of the CFG and checks
that every vertex within a task is dominated by the entry vertex of the task.
If this dominance condition is unmet for a CFG node, the compiler places an
additional task boundary at that node and re-evaluates the condition, repeatedly
inserting a boundary at each node for which the dominance condition is unmet.
Figure~\ref{fig:compiler_overview} illustrates boundary placement: the first
three vertices make up one task and the remainder become a second task.

@@

After forming tasks, variables that are live beyond the end of the task become
protected, which requires renaming those variables and updating references to
renamed variable names. For example, Figure~\ref{fig:compiler_overview} shows a
boundary placed within the live range of {\tt b}; {\tt b} becomes protected and
accesses to {\tt b} should be redirected to {\tt p\_b} (as shown in
Figure~\ref{fig:compiler_overview}). {\tt a} is task-local and need not be
protected, but only declared in the task. Constants and read-only variable are
never protected. When writes in multiple, newly formed tasks may update the
same location (i.e., at a PHI node in the CFG), \sys re-writes the accesses to
refer to the same protected, global name. Subsequent reads simply read the
protected variable.

>>>>

We studied \sys's effectiveness for systems with automatically generated task
boundaries by developing a simple compiler prototype that transforms an
arbitrary C into a program with tasks. The compiler is representative of,
although not a direct implementation of, a system that assists in writing
task-based programs by automatically inserting tasks (e.g., inspired by
Ratchet~\cite{ratchet}). Translating C code to use tasks requires identifying program points to use as task boundaries, defining a set of task functions that begin and end at those boundaries, ensuring that all existing control-flow paths are either contained within a task or are made explicit with inter-task control-flow, and identifying and instrumenting accesses to task-shared variables that need to be protected by \sys at runtime. Our simple prototype, implemented using LLVM, analyzes the control-flow graph (CFG) and uses a rule-based heuristic to determine at which points to place a task boundary. After identifying where to put task
boundaries, the \sys compiler transforms the imperative code into tasks,
instrumenting protected memory accesses.

@@

\sys analyzes a program to determine where to put task boundaries. 
%
A task is single-entry, multiple-exit part of the CFG that \sys compiler excises
into a new function.  
%
There are several factors determining efficient task execution. A task decomposition should minimize the number of variables that are accessed by multiple tasks and, consequently, must be protected by \sys's paging mechanism. 
%
Different tasks should be approximately the same size to make reasoning about
coalescing simple by eliminating the need to consider variable run time costs. 
%
To minimize task boundary overheads (e.g., commit), a task decomposition should
maximize task size, which minimizes the frequency of task transitions.

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/compiler.pdf}
	\caption{\sys compiles code to use tasks.}
	\label{fig:compiler_overview}
\end{figure}

The task boundary identification analysis processes a standard C program (as
shown in Figure~\ref{fig:compiler_overview}) and produces a CFG with some
vertices marked as {\em task boundaries}. In our simple
reference prototype, we implemented inter-procedural analysis by inlining all
functions called.  We opted not to support recursion, which is uncommon in
embedded code. We leave to future work better approaches to inter-procedural
analysis and task transformation.

@@

\textbf{Estimated Run Time.} \sys assigns each basic block in the CFG a
{\em cost}, i.e. an estimate of the run time of the block. Precisely
modelling the run time and energy cost of arbitrary C code is difficult and
beyond the scope of our simple compiler prototype. Instead, our prototype counts
LLVM intermediate representation (IR) instructions and assigns a higher cost to
a block with more operations, assuming that the block runs for a longer period
of time and consumes more energy. To make the cost estimate useful, our compiler prototype estimates the maximum cost that the device's energy buffering capacitor supports. \sys does this by measuring the number of arithmetic operations that can execute without a power interruption, starting from a full capacitor. This maximum number of operations is a limit on the cost of a task---any longer task risks not executing to completion.

\textbf{Live Variable Cost.} Each CFG edge is associated with a set of
variables that {\em may} be {\em alive} across the edge, i.e., variables
assigned before the edge and may be read after the edge. Variables that live
across an edge must be protected by \sys's memory virtualization mechanism when
\sys places a task boundary on the edge. \sys heuristically minimizes variable
virtualization overhead by placing boundaries to minimize the number of
live variables that span task boundaries.

\textbf{Loop Cost.} Loops are a challenge for \sys because the analysis
must account for the cost of multiple iterations of the loop body. \sys assumes
that most loops are annotated with a static loop bound. For loops that do not
have a static bound, \sys conservatively assumes the loop will be too large to
fit within a single task. Assuming loop bounds is a reasonable simplification:
while not always the case, long loops in embedded code are often statically
bounded, especially when code aims to meet strict resource requirements.

@@

After placing task boundaries, the compiler must ensure that all tasks are
single-entry. To do so, \sys computes the dominator tree of the CFG and checks
that every vertex within a task is dominated by the entry vertex of the task.
If this condition is unmet for a CFG node, the compiler places an
additional task boundary at that node and re-evaluates the condition, repeatedly
inserting a boundary at each node for which the dominance condition is unmet.
Figure~\ref{fig:compiler_overview} illustrates boundary placement: the first
three vertices make up one task and the remainder become a second task.

@@

After forming tasks, variables that are live beyond the end of the task become
protected, which requires renaming those variables and updating references to
renamed variable names. For example, Figure~\ref{fig:compiler_overview} shows a
boundary placed within the live range of {\tt b}; {\tt b} becomes protected and
accesses to {\tt b} should be redirected to {\tt p\_b} (as shown in
Figure~\ref{fig:compiler_overview}). {\tt a} is task-local and need not be
protected, but only declared in the task. Constants and read-only variable are
never protected. When writes in multiple, newly formed tasks may update the
same location, i.e., at a PHI node in the CFG, \sys re-writes the accesses to
refer to the same protected, global name. Subsequent reads simply read the
protected variable.
